---
title: "Time Series Analysis, Case 6"
author: "Andew, Carmen, Jimmy, Naresh"
date: "23 January 2016"
output: 
  pdf_document: 
    latex_engine: xelatex
---

We are going to analyze six simulated time series, and in all of them we have to decide,
in a visual and graphical way, if they are or not Covariance-Stationary (CS), White Noise
(WN), Strict White Noise (SWN) or Gaussian White Noise (GWN). We will also analyze
the real time series for the daily Brent Dated spot prices and its corresponding returns.

#Loading data
First, we load the two datasets
```{r}
getwd()
setwd("/Users/nareshshah/downloads")
library(ggplot2)
library(gridExtra)
real = read.csv("Session6real.csv", sep = "",dec = ",")
names(real) = "price"
sim = read.csv("Session6sim.csv", sep = ";",dec = ",")
```


##Analysis of the Real Data
We need to add a difference value that look at the difference between the price and the previous day. To do this we create a lag of t-1 and then calculate the different.
```{r}
library(dplyr)
real$lag = lag(real$price,1)
real$diff = real$price - real$lag
ts.plot(real$diff)
```


###Step 1. Plot the series and compute basic statistics
First we compute the mean, standard deviation and the variance of the time series. We then plot the series and see the data is not stationary. The variance is different across the time series so can assume it is not stationary (is this right guys?). 
```{r}
summary(real$price)
sd(real$price)
var(real$price)
ts.plot(real$price, main = "Time Series Plot of Daily Brent Spot Prices ($/bbl)")
ts.plot(real$diff, main = "Time Series Plot of Difference in Daily Brent Spot Prices ($/bbl)")
```

###Step 2. Plot the acf and pacf for the series
We now plot the Auto Correlation Function and the Partial AutoCorrelation Function
```{r}
acf(real$price, main = "Auto Correlation Function Plot of Daily Brent Spot Prices ($/bbl)")
pacf(real$price,type = "partial", main = "Partial Auto Correlation Function Plot of Daily Brent Spot Prices ($/bbl)")
#Since the auto-correlation is relatively low, it is unlikely that this is white noise.
```


###Step 3. Test Normality

1. A stochastic process is covariance stationary (or weak stationary) if
  a. E(Yt) = µ, for all t
  b. Var(Yt) = σy2, for all t
  c. Cov(Yt,Yt-h) = ϒ(h), for all t and h
2. A process is white noise if it is
  a. Covariance Stationary
  b. Uncorrelated
  c. Zero mean
3. A process is strict white noise if it is
  a. Covariance Stationary
  b. Zero mean
  c. Independent and identically distributed (iid)

ts.plot(log(real), main = "Time Series Plot of Daily Brent Spot Prices ($/bbl)")

#Step 4: Checking simulated Data statistics
```{r}
summary(sim)
mean(sim)
p1 = qplot(sim$Serie1)
p2 = qplot(sim$Serie2)
p3 = qplot(sim$Serie3)
p4 = qplot(sim$Serie4)
p5 = qplot(sim$Serie5)
p6 = qplot(sim$Serie6)

grid.arrange(p1,p2,p3,p4,p5,p6, nrow = 3)

M = cor(sim)
#corrplot(M,method = 'number')
#corrplot(M,method = 'square')
#mean = summarize(sim,mean(sim,na.rm="TRUE"))

```
#Step 5: Checking Series 1
```{r}
mean1 = mean(sim$Serie1,na.rm = "TRUE")
sd1 = sd(sim$Serie1,na.rm="TRUE")
var1 = var(sim$Serie1,na.rm="TRUE")
ts.plot(sim$Serie1[0:200])
Series1 = na.omit(sim$Serie1)
acf(Series1)
#Since most auto-correlation values are close to 0, this might indicate white noise.
#Mean is not close to 0, not white noise.
# Total of 3 auto-correlation points above critical values according to the chart. 3/200 does not cross the 95% of points under threshold limit. 
# Mean remains somewhat constant across time
mean(Series1[0:10])
mean(Series1[0:100])
mean(Series1[100:200])
#Here we can use a hypothesis test to determine if mean is somewhat centered around 2, but it is quite obvious from the graph and the above three tests.
sd(Series1[0:10])
sd(Series1[0:100])
sd(Series1[100:200])
#Here standard Deviation and variance seem to decrease with time. Therefore, this series is not stationary either.
```
#Step 6: Checking Series 2
```{r}
mean2 = mean(sim$Serie2,na.rm = "TRUE")
sd2 = sd(sim$Serie2,na.rm="TRUE")
var2 = var(sim$Serie2,na.rm="TRUE")
ts.plot(sim$Serie2[0:200])
Series2 = na.omit(sim$Serie2)
acf(Series2)
#Since most auto-correlation values are close to 0, this might indicate white noise.
#Mean is close to 0, Mean = 0.13
# Total of 2 auto-correlation points above critical values according to the chart. 2/200 does not cross the 95% of points under threshold limit. 
# Mean decreases across time
mean(Series2[0:10])
mean(Series2[0:100])
mean(Series2[100:200])
#Here we can use a hypothesis test to determine if mean is somewhat centered around 0, but it is quite obvious from the graph and the above three tests.
sd(Series2[0:10])
sd(Series2[0:100])
sd(Series2[100:200])
#Here standard Deviation and variance seem to be roughly constant. Therefore, this series is not covariance stationary either.
lag2 = lag(Series2)
cov(Series2,lag2,use = 'complete')
cov(Series2[0:10],lag2[0:10],use = 'complete')
cov(Series2[0:100],lag2[0:100],use = 'complete')
cov(Series2[100:200],lag2[100:200],use = 'complete')
#This tells us that covariance is roughly constant.
shapiro.test(Series2)
#High p-value means the white noise is normal and therefore gaussian. (SWN,GWN)

```

#Step 7: Checking Series 3
```{r}
mean3 = mean(sim$Serie3,na.rm = "TRUE")
sd3 = sd(sim$Serie3,na.rm="TRUE")
var3 = var(sim$Serie3,na.rm="TRUE")
ts.plot(sim$Serie3[0:200])
Series3 = na.omit(sim$Serie3)
acf(Series3)
#Since most auto-correlation values are close to 0, this might indicate white noise.
#Mean is close to 0, Mean = -0.10494
# Total of 2 auto-correlation points above critical values according to the chart. 2/200 does not cross the 95% of points under threshold limit. 
# Mean increases across time
mean(Series3[0:10])
mean(Series3[0:100])
mean(Series3[100:200])
#Here we can use a hypothesis test to determine if mean is somewhat centered around 0, but it is quite obvious from the graph and the above three tests.
sd(Series3[0:10])
sd(Series3[0:100])
sd(Series3[100:200])
#Here standard Deviation and variance seem to decrease with time. Therefore, this series is not covariance stationary either.
lag3 = lag(Series3)
cov(Series3,lag3,use = 'complete')
cov(Series3[0:10],lag3[0:10],use = 'complete')
cov(Series3[0:100],lag3[0:100],use = 'complete')
cov(Series3[100:200],lag3[100:200],use = 'complete')
#This tells us that covariance is dependent on time and constantly decreasing based on time. Which makes this not white noise. Removing the damping factor would probably convert this into white noise.


```

#Step 8: Checking Series 4
```{r}
mean4 = mean(sim$Serie4,na.rm = "TRUE")
sd4 = sd(sim$Serie4,na.rm="TRUE")
var4 = var(sim$Serie4,na.rm="TRUE")
ts.plot(sim$Serie4[0:200])
Series4 = na.omit(sim$Serie4)
acf(Series4)
#Since most auto-correlation values are close to 0, this might indicate white noise.
#Mean is close to 0, Mean is -0.080695
# Total of 2 auto-correlation points above critical values according to the chart. 2/200 does not cross the 95% of points under threshold limit. 
# Mean is roughly constant
mean(Series4[0:10])
mean(Series4[0:100])
mean(Series4[100:200])
#Here we can use a hypothesis test to determine if mean is somewhat centered around 0, but it is quite obvious from the graph and the above three tests.
sd(Series4[0:10])
sd(Series4[0:100])
sd(Series4[100:200])
#Here standard Deviation and variance seem to roughly constant. Therefore, this series is covariance stationary.
lag4 = lag(Series4)
cov(Series4,lag4,use = 'complete')
cov(Series4[0:10],lag4[0:10],use = 'complete')
cov(Series4[0:100],lag4[0:100],use = 'complete')
cov(Series4[100:200],lag4[100:200],use = 'complete')
#This tells us that covariance is independent of time and almost constant.
#Testing for normality
shapiro.test(Series4)
#It seems like Series 4 is a normally distributed white noise (Gaussian White noise and strictly white noise) (SWN,GWN)

```

#Step 9: Checking Series 5
```{r}
mean5 = mean(sim$Serie5,na.rm = "TRUE")
sd5 = sd(sim$Serie5,na.rm="TRUE")
var5 = var(sim$Serie5,na.rm="TRUE")
ts.plot(sim$Serie5[0:200])
Series5 = na.omit(sim$Serie5)
acf(Series5)
#Since most auto-correlation values are not close to 0, this indicates a definite time series pattern.
# Mean is far from 0
# Mean decreases across time
mean(Series5[0:10])
mean(Series5[0:100])
mean(Series5[100:200])
#Here we can use a hypothesis test to determine if mean is somewhat centered around 0, but it is quite obvious from the graph and the above three tests.
sd(Series5[0:10])
sd(Series5[0:100])
sd(Series5[100:200])
#Here standard Deviation and variance seem to increase with time. Therefore, this series is not covariance stationary either.
lag2 = lag(Series2)
cov(Series2,lag2,use = 'complete')
cov(Series2[0:10],lag2[0:10],use = 'complete')
cov(Series2[0:100],lag2[0:100],use = 'complete')
cov(Series2[100:200],lag2[100:200],use = 'complete')
#This tells us that covariance is independent of time and roughly constant.
```

#Step 10: Checking Series 6
```{r}
mean6 = mean(sim$Serie6,na.rm = "TRUE")
sd6 = sd(sim$Serie6,na.rm="TRUE")
var6 = var(sim$Serie6,na.rm="TRUE")
ts.plot(sim$Serie6)
Series6 = na.omit(sim$Serie6)
acf(Series6)
#Since most auto-correlation values are close to 0, this might indicate white noise.
#Mean is close to 0, Mean is -0.0002929
# Mean roughly constant
mean(Series6[0:10])
mean(Series6[0:1000])
mean(Series6[1000:200])
#Here we can use a hypothesis test to determine if mean is somewhat centered around 0, but it is quite obvious from the graph and the above three tests.
sd(Series6[0:10])
sd(Series6[0:1000])
sd(Series6[1000:2000])
#Here standard Deviation and variance seem to be roughly constant/decrease with time. Therefore, this series is perhaps covariance stationary.
lag6 = lag(Series6)
cov(Series6,lag6,use = 'complete')
cov(Series6[0:10],lag6[0:10],use = 'complete')
cov(Series6[0:1000],lag6[0:1000],use = 'complete')
cov(Series6[1000:2000],lag6[1000:2000],use = 'complete')
#This tells us that covariance is independent of time and roughly constant.
shapiro.test(Series6)
# Not a normally distributed but definitely white noise.(WN)

```

